---
title: "Project 5"
author: "Justin Williams"
date: "`r Sys.Date()`"
output: html_document
---

---
title: "Project 5"
author: "Justin Williams"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#import libraries and data
library(tidyverse)
library(dplyr)
library(reshape2)
library(ggplot2)
library(kableExtra)
hrdata_df <- read.csv("/Users/justinwilliams/Code/9050advresearch/Project 5/HRData.csv")
```

# 1. Run a simple regression analysis in which you predict PerfScoreID from EmpSatisfaction. Provide a summary of this analysis like what you would find in a journal article. Be sure to provide a table of results, a plot of the regression line, AND a written summary of the results in your response.
```{r}
# Simple regression analysis
simple_model <- lm(PerfScoreID ~ EmpSatisfaction, data = hrdata_df)

# Model summary
simple_model_summary <- summary(simple_model)
```
```{r}
# Model results data frame
simple_results_df <- data.frame(
  Predictor = rownames(coef(simple_model_summary)),
  Estimate = coef(simple_model_summary)[, "Estimate"],
  Std_Error = coef(simple_model_summary)[, "Std. Error"],
  t_value = coef(simple_model_summary)[, "t value"],
  P_value = coef(simple_model_summary)[, "Pr(>|t|)"]
)

print(simple_results_df)
```
```{r}
#print results in a table
print("Simple Regression Results Data Frame:")
kable(simple_results_df, format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(0, bold = TRUE, color = "white", background = "orange") %>%
  column_spec(1, bold = TRUE, color = "purple")
```
```{r}
# Plot of the regression line
ggplot(hrdata_df, aes(x = EmpSatisfaction, y = PerfScoreID)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Regression of PerfScoreID on EmpSatisfaction",
       x = "EmpSatisfaction",
       y = "PerfScoreID")
```

## Summary
The regression analysis demonstrates that EmpSatisfaction significantly predicts PerfScoreID. The slope shows that for every one-unit increase in EmpSatisfaction, PerfScoreID increases by 0.196. The intercept suggests that when EmpSatisfaction is 0, the predicted PerfScoreID is 2.215. 9.22% of the variability in PerfScoreID is accounted for by EmpSatisfaction.

## a. Is the regression weight statistically significant?
Yes, p < 0.001.

## b. What is the regression equation?
PerfScoreID = 2.215 + 0.196 * EmpSatisfaction

## c. Provide an interpretation of the slope in terms of the variables involved.
For every one-unit increase in EmpSatisfaction, PerfScoreID increases by 0.196.

## d. Interpret the y-intercept.
When EmpSatisfaction is 0, the predicted PerfScoreID is 2.215.

## e. What is the predicted PerfScoreID for someone with an average level of EmpSatisfaction?
The predicted PerfScoreID would be calculated as Y = 2.215 + 0.196 * the mean EmpSatisfaction.

## f. What percentage of variance in PerfScoreID is explained by EmpSatisfaction?
EmpSatisfaction explains 9.22% of the variance in PerfScoreID.