---
title: "Project 4"
author: "Justin Williams"
date: "`r Sys.Date()`"
output: pdf_document
---

---
title: "Project 2"
author: "Justin Williams"
date: "`r Sys.Date()`"
output: html_document
---
Packages, data, etc.
```{r}
#import libraries and data
library(tidyverse)
library(dplyr)
library(reshape2)
library(ggplot2)
hrdata_df <- read.csv("/Users/justinwilliams/Code/9050advresearch/Project 4/HRData.csv")
```

# 1.Report the full covariance and correlation matrix for the following variables: GenderID, PerfScoreID, Salary, Age, EngagementSurvey, EmpSatisfaction, and Absences.  
```{r}
#select the columns

selected_columns <- hrdata_df[, c("GenderID", "PerfScoreID", "Salary", "Age", "EngagementSurvey", "EmpSatisfaction", "Absences")]
#covariance matrix
cov_matrix <- cov(selected_columns)
print("Covariance Matrix:")
print(cov_matrix)

#correlation matrix
cor_matrix <- cor(selected_columns)
print("Correlation Matrix:")
print(cor_matrix)
```

# 2. Visualize the correlations you have computed. There are a few different ways to do this (including scatterplots, heat maps, etc.). Use the visualization method that you think best conveys the relations. 
```{r}
#converting matrices to long format for ggplot
cov_melt <- melt(cov_matrix)
cor_melt <- melt(cor_matrix)

#heatmap for covariance matrix
ggplot(data = cov_melt, aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_gradient2(low = "purple", high = "orange", mid = "white", midpoint = 0) +
    theme_minimal() +
    labs(title = "Covariance Matrix Heatmap", x = "", y = "")

# heatmap for correlation matrix
ggplot(data = cor_melt, aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_gradient2(low = "purple", high = "orange", mid = "white", midpoint = 0) +
    theme_minimal() +
    labs(title = "Correlation Matrix Heatmap", x = "", y = "")
```
# 3. Provide a summary of what you see in these results. Explain the nature of the bivariate relations (or lack thereof) in these data. Write this summary like you might see reported in a journal article. 
 
## Create a subset of the data in which you only include employees who are from the Production Department. Also, create a subset of everyone NOT in Production.  
 
# 4. Run correlations separately for each of these two subsamples. Report the results for each and provide an interpretation of what you see as you did in Questions 2 and 3. Are the correlations the same across the subsamples and/or from the entire sample? I’m not asking for a test of statistical significance, but your own interpretation based on the values you compute. 